#Reproducible Research - Peer Assessment 1

## Loading and preprocessing the data
1. Load the data

```{r, echo=TRUE}
if(!file.exists('activity.csv')){
    unzip('activity.zip')
}
data <- read.csv('activity.csv')
```
2. Convert date to date data type

```{r, echo=TRUE}
data$date <- as.Date(data$date) 
```

## What is mean total number of steps taken per day?
1. Calculate the total number of steps taken per day
```{r, echo=TRUE}
totalStepsPerDay <- aggregate(steps ~ date, data = data, sum, na.rm = TRUE)
```
2. Make a histogram of the total number of steps taken each day
```{r, echo=TRUE}
hist(totalStepsPerDay$steps, main = "Total steps taken per day", xlab = "number of steps", breaks=13)
```
  
3. Calculate and report the mean and median of the total number of steps taken per day
```{r, echo=TRUE}
meanStepsPerDay <- mean(totalStepsPerDay$steps)
meanStepsPerDay
medianStepsPerDay <- median(totalStepsPerDay$steps)
medianStepsPerDay
```


## What is the average daily activity pattern?

1. Calculate average steps for each of 5-minute interval during a 24-hour period

```{r, echo=TRUE}
interval.mean.steps <- aggregate(steps ~ interval, data = data, mean, na.rm = TRUE)
```
2. Make a time series plot of the 5-minute interval

```{r, echo=TRUE}
plot(interval.mean.steps$interval, interval.mean.steps$steps, type = "l", col="blue", 
     main="Average number of steps taken", xlab=" 5-minute interval", 
     ylab="Average number of steps")
```


2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r, echo=TRUE}
interval.mean.steps[interval.mean.steps$steps == max(interval.mean.steps$steps), ]
```

## Imputing missing values

1. Calculate and report the total number of missing values in the dataset
```{r, echo=TRUE}
numMissingValues <- length(which(is.na(data$steps)))

```

2. Devise a strategy for filling in all of the missing values in the dataset.

To fill the missing values in the dataset, here I use the mean for that 5-minute interval

3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r, echo=TRUE}
library(Hmisc)
dataImputed <- data
dataImputed$steps <- impute(data$steps, fun=mean)
```
4. Make a histogram of the total number of steps taken each day
```{r, echo=TRUE}
stepsByDayImputed <- tapply(dataImputed$steps, dataImputed$date, sum)

hist(stepsByDayImputed, main = "Total number of steps taken per day", xlab = "Steps", ylim =c(0, 25),breaks=13)
```

5. Calculate and report the mean and median of the total number of steps taken per day from imputed data
```{r, echo=TRUE}
meanStepsPerDayImputed <- mean(stepsByDayImputed)
meanStepsPerDayImputed
medianStepsPerDayImputed <- median(stepsByDayImputed)
medianStepsPerDayImputed
```

## Are there differences in activity patterns between weekdays and weekends?
1.Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r, echo=TRUE}
wd <- weekdays (data$date)
data$dateType <- factor (wd, levels <- c ('weekday', 'weekend'))
data$dateType[] <- 'weekday'
data$dateType[wd %in% c ('Saturday', 'Sunday')] <- 'weekend'
```

2. Make a panel plot containing a time series plot

```{r, echo=TRUE}
library(plyr)
dataImputed <- ddply (data,  .(interval, dateType), summarize, 
              stepMean = mean(steps, na.rm=TRUE))

xyplot(stepMean ~ interval | dateType, 
        data=dataImputed,
        type='l', layout=c (1, 2))
```